---
title: "Week 11"
author: "Raj, Ponnam"
date: March 05th, 2022
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



# a. Plot the data from each dataset using a scatter plot.

```{r}
library('ggplot2')
library('class')

# working directory path to access binary-classifier-data.CSV file from local drive
setwd("/Users/rajponnam/Documents/ds_course/dsc520/data")

# Read the files `binary-classifier-data.csv` and trinary-classifier-data.csv to `binary_df` and 'trinary_df'
binary_df <- read.csv("binary-classifier-data.csv")
trinary_df <- read.csv("trinary-classifier-data.csv")

temp_binary_df <- binary_df
temp_binary_df$label <- as.factor(temp_binary_df$label)
temp_trinary_df <- trinary_df
temp_trinary_df$label <- as.factor(temp_trinary_df$label)

# Plot the data dataset using a scatter plot for both files
ggplot(temp_binary_df, aes(x=x, y=y, color=label)) + geom_point()
ggplot(temp_trinary_df, aes(x=x, y=y, color=label)) + geom_point()

```

# b.In this problem, you will determine which points are nearest by calculating the Euclidean distance between two points.
```{r}

library ('TSdist')

EuclideanDistance(binary_df$x,binary_df$y)

EuclideanDistance(trinary_df$x,trinary_df$y)

```
# Fit a k nearest neighbors model for each dataset for k=3, k=5, k=10, k=15, k=20, and k=25. Compute the accuracy of the resulting models for each value of k. Plot the results in a graph where the x-axis is the different values of k and the y-axis is the accuracy of the model.

knn for binary
```{r}
# Split data to use 80% of data to train the model and 20% of data to test the model
data_split_binary <- sample(1:nrow(binary_df), 0.8 * nrow(binary_df))
train_binary <- binary_df[data_split_binary,]
test_binary <- binary_df[-data_split_binary,]

binary_glm <- glm(label ~ x + y, data=binary_df, family = binomial)

# extract 1st column of train dataset because it will be used as 'cl' argument in knn function.
target_category <- binary_df[data_split_binary,1]

# extract 1st column if test dataset to measure the accuracy
test_category <- binary_df[-data_split_binary,1]

k_values<- c(3, 5, 10, 15, 20, 25)
Accuracy <- NULL
for (i in 1:length(k_values))
{
  test_pred <- knn(train_binary,test_binary,cl=target_category,k=k_values[i])
  confmatrix <- table(test_category,test_pred)
  accuracy <- (confmatrix[[1,1]] + confmatrix[[2,2]]) / sum(confmatrix)
  Accuracy <- c(Accuracy, round((accuracy * 100), digits=2))
}
binary_results_df <- data.frame(k_values, Accuracy)
binary_results_df
ggplot(binary_results_df, aes(x=k_values, y=Accuracy)) + geom_point() + geom_line(colour="red")

```
knn for trinary 
```{r}
# Split data to use 80% of data to train the model and 20% of data to test the model
data_split_trinary <- sample(1:nrow(trinary_df), 0.8 * nrow(trinary_df))
train_trinary <- trinary_df[data_split_trinary,]
test_trinary <- trinary_df[-data_split_trinary,]

trinary_glm <- glm(label ~ x + y, data=trinary_df, family = poisson)

# extract 1st column of train dataset because it will be used as 'cl' argument in knn function.
target_category <- trinary_df[data_split_trinary,1]

# extract 1st column if test dataset to measure the accuracy
test_category <- trinary_df[-data_split_trinary,1]

k_values<- c(3, 5, 10, 15, 20, 25)
Accuracy <- NULL
for (i in 1:length(k_values))
{
  test_pred <- knn(train_trinary,test_trinary,cl=target_category,k=k_values[i])
  confmatrix <- table(test_category,test_pred)
  accuracy <- (confmatrix[[1,1]] + confmatrix[[2,2]]) / sum(confmatrix)
  Accuracy <- c(Accuracy, round((accuracy * 100), digits=2))
}
trinary_results_df <- data.frame(k_values, Accuracy)
trinary_results_df
ggplot(trinary_results_df, aes(x=k_values, y=Accuracy)) + geom_point() + geom_line(colour="red")

```


# C. Looking back at the plots of the data, do you think a linear classifier would work well on these datasets?
Linear classifier will not work for both binary and trinary datasets because data points are not grouped around a line. Data points are spread all over the graph.  
